Determinants of Soccer Players' Market Values in Major European League(s)

### 2024-12-04 paper/paper.qmd
Paper final update

---
title: "Determinants of Soccer Players' Market Values in Major European League(s)"
subtitle: "Goals, Assists, and Age Dominate, While League-Specific Patterns Add Complexity"
author: 
  - John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Player_Market_Value_Analysis"
date: today
date-format: long
abstract: "The global soccer transfer market is shaped by diverse factors influencing player valuations across leagues. This study analyzes determinants of market value in five major European leagues using data from the 2023/24 season. Goals and assists consistently emerge as the strongest predictors, with the Premier League valuing goals highest and the Bundesliga emphasizing assists, while age negatively affects value across all leagues, particularly in La Liga. These insights highlight universal trends and league-specific valuation dynamics, offering actionable guidance for clubs, agents, and analysts."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The global soccer transfer market is a high-stakes, multi-billion-Euro industry where decisions on player valuations can significantly influence team-building strategies, investment decisions, and negotiations. Player market values are determined by a mix of tangible performance metrics, such as goals scored or assists provided, and intangible factors like potential, reputation, and team context. While certain predictors are universally recognized, such as offensive contributions and age, the relative importance of these factors varies widely across leagues with differing financial capabilities, tactical preferences, and competitive levels. Understanding the dynamics behind these variations is crucial for clubs, agents, and analysts aiming to optimize their strategies in this complex market.

This study examines the determinants of soccer players’ market values across five major European leagues: Premier League(England), La Liga (Spain), Serie A (Italy), Bundesliga (Germany), and Ligue 1 (France). These leagues not only dominate the global football landscape but also present unique economic and competitive ecosystems, making them ideal for a comparative analysis. Using player data from the 2023/24 season, sourced from platforms such as @citeTransfermarkt, @citeStathead, @citeFIFA, and @citeFootballDatabase, this research addresses a critical gap: how do the drivers of player market value differ across these leagues?

@@

The results reveal both universal and league-specific patterns in player valuation. Offensive contributions, such as goals and assists, consistently emerged as significant predictors, with goals having the strongest influence in the Premier League and assists being most impactful in the Bundesliga. Age also played a critical role, with younger players commanding higher valuations, particularly in La Liga. Contextual variables like club and national team rankings demonstrated notable variation, reflecting the differing emphasis on team prestige across leagues.

This study contributes to the literature on soccer economics by uncovering how market dynamics vary across Europe’s top leagues. For clubs, the findings provide actionable insights for tailoring recruitment and investment strategies. For agents, the results highlight the most marketable player attributes for negotiations. Methodologically, the integration of advanced statistical and programming tools offers a replicable framework for analyzing market dynamics in soccer and other sports.

The paper is structured as follows: @sec-data describes the dataset and variable construction processes, emphasizing the integration of data from multiple sources. @sec-model details the modeling approach and validation techniques. @sec-res presents the empirical results, identifying key differences across leagues. @sec-disc discusses the implications of these findings, highlights the limitations of the study, and outlines future research directions. @sec-appendix provides an in-depth exploration of data cleaning methodologies, comprehensive documentation of observational data collection processes, datasheets for each dataset, and full model results. By bridging methodological rigor with actionable insights, this study advances our understanding of the determinants of soccer players’ market values.

@@

The dataset was created by merging multiple data sources using player names as unique identifiers. **Transfermarkt.com** provided player market values in Euros, which serve as the dependent variable in this study. Performance metrics such as goals and assists were sourced from **Stathead.com**, a platform that aggregates comprehensive statistics on football players. To ensure consistency and comparability, rankings were applied to national teams and clubs. National team rankings (1 to 210) were extracted from **FIFA.com**, while club rankings were sourced from **Footballdatabase.com**, reflecting both domestic and international performances.

After merging these sources, high-level cleaning was conducted to address missing data and standardize variable formats, and the players who are classified as goalkeepers are dropped. The final cleaned datasets for each league include seven variables, as outlined below:

@@

The England dataset provides a comprehensive view of player characteristics in the English Premier League. As illustrated in @fig-emp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower market value brackets (below €50 million). This skewness reflects a superstar effect, where a small subset of players, often international stars or highly sought-after talent, command disproportionately higher valuations.

@fig-efp explores the distributions of key predictor variables, including age, goals, assists, national team ranking, club ranking, and minutes played. These distributions reveal important patterns in the dataset:

@@

# Reshape data for faceting (excluding market_value)
data_long <- england_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "age",
      "goals",
      "assists",
      "national_team_ranking",
      "club_ranking",
      "minutes_played"
    ))
  )

@@

# Summarize numeric variables
numeric_summary <- england_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The France dataset offers insights into player profiles from Ligue 1, one of Europe’s top football leagues. Similar to other leagues, the distribution of market values in the French dataset, as depicted in @fig-fmp, is heavily right-skewed. Most players have market values below €50 million, with only a few elite players commanding significantly higher valuations. This reflects the trend where top-tier talents, often representing national teams or excelling in international competitions, are valued much higher than their peers.

@@

- **Age**: The age distribution in Ligue 1 peaks around 24 years, with a majority of players between 20 and 30 years old. This highlights the league's balance between developing young talent and retaining experienced professionals. The maximum age in the dataset is 39 years, underscoring the presence of veteran players in the league.
- **Goals and Assists**: The distributions for goals and assists are highly right-skewed, with most players contributing fewer than 5 goals or assists in a season. This indicates that offensive production is concentrated among a small group of players, typically forwards or attacking midfielders.
- **Club Ranking**: Club rankings span a wide range, from 11 to 568, reflecting the diversity in club performance levels within the league.
- **National Team Ranking**: National team rankings also show significant variation, with players representing countries ranked in the top 10 as well as those from lower-ranked national teams. This diversity emphasizes Ligue 1's role as a platform for both elite and emerging talent.
- **Minutes Played**: The distribution of minutes played indicates that a significant proportion of players log substantial game time (over 1,000 minutes), with some exceeding 3,000 minutes, indicating their critical role in team strategies.

@@

@tbl-sf provides a comprehensive statistical summary of the numeric variables in the dataset. The mean market value stands at €7.98 million, significantly lower than the Premier League, indicating the varying financial dynamics across leagues. The average age is 23.91 years, reinforcing the league's focus on younger talent. Other variables such as goals and assists demonstrate low means relative to their maximum values, reflecting the rarity of exceptional performances.

@@

# Reshape data for faceting and specify the order of variables
data_long <- france_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

@@

# Summarize numeric variables for France
numeric_summary_france <- france_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

@fig-gfp explores the distributions of predictor variables, shedding light on key patterns:

@@

The position distribution, depicted in @fig-gpp, reveals that defenders (DF) comprise the largest share of players (31.7%), followed by midfielders (MF) and forward-midfield hybrids (FWMF). Pure forwards (FW) form 9.6% of the players, while hybrid positions like FWDF make smaller contributions. This composition underscores the league's focus on tactical balance, with strong emphasis on midfield and defensive strategies.

@@

# Reshape data for faceting (excluding market_value)
data_long <- germany_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

@@

# Summarize numeric variables for Germany
numeric_summary_germany <- germany_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The Italy dataset provides an insightful analysis of player characteristics within the Serie A league. As shown in Figure @fig-imp, the market value distribution is right-skewed, with a majority of players concentrated in lower valuation brackets (below €30 million). This distribution is characteristic of the league, where the presence of a few superstar players significantly elevates the upper tail.

@@

# Reshape data for faceting and specify the order of variables
data_long <- italy_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

@@

# Summarize numeric variables for Italy
numeric_summary_italy <- italy_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The Spain dataset offers a comprehensive analysis of player characteristics in La Liga. As illustrated in @fig-smp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower valuation brackets (below €30 million). This trend reflects the league's structure, where a few superstar players dominate the upper tail of the distribution, commanding significantly higher valuations.

@fig-sfp reveals the distributions of key predictor variables:

@@

@fig-spp illustrates the percentage distribution of player positions. Defenders (DF) comprise the largest group (32.3%), followed by midfielders (MF) at 24.4%. Hybrid roles such as forward-midfield (FWMF) and forward (FW) are well-represented, reflecting La Liga’s tactical diversity and emphasis on attacking football. Smaller proportions of hybrid defensive positions (FWDF, MFDF) further highlight the league's strategic nuances.

@@

# Reshape data for faceting and specify the order of variables
data_long <- spain_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

@@

# Summarize numeric variables for Spain
numeric_summary_spain <- spain_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The construction of variables in this study ensures their relevance and accuracy. Player market values were obtained from Transfermarkt, a widely regarded source for estimating market valuations. These values are based on a combination of player performance, market demand, and subjective expert evaluation, making them a holistic representation of player worth. Performance metrics, such as goals and assists, were sourced from Stathead, a trusted repository of sports data.

National team and club rankings were quantified to enable comparison across leagues. National team rankings were sourced from FIFA’s official ranking system, which reflects international match results, strength of opposition, and match importance. Club rankings were sourced from Footballdatabase, which evaluates team performance using a consistent point-based methodology.

@@

The selected dataset is uniquely suited for this analysis due to its comprehensive scope and integration of market, performance, and contextual data. While alternative sources such as Opta or WyScout provide more granular match-level data, their focus on in-game metrics (e.g., pass accuracy, xG) lacks the broader valuation context provided by Transfermarkt. Moreover, FIFA rankings and Footballdatabase rankings are well-established benchmarks for quantifying national and club performance, making them preferable over less transparent systems.

@@

In this study, we constructed a linear regression model for each of the five major European football leagues to analyze the determinants of player market value. The dependent variable, market value $(\text{Market Value}_i)$, was scaled to millions of Euros for easier interpretability. The model incorporates a comprehensive set of predictors, including player performance metrics, contextual factors, and positional variables, to capture the multi-dimensional aspects influencing market valuation.

@@

The models were implemented using the R programming language, specifically leveraging the `lm()` function for linear regression. For each league, data were split into training and testing sets to validate model performance. A stratified train-test split (80% training, 20% testing) was performed using the `caret` package to ensure balanced representation of market value distributions across the sets.

@@

Age negatively affects market value across all leagues, reflecting the premium placed on youth. @fig-ae reveals that La Liga (Spain) places the strongest emphasis on younger players, with a coefficient of -1.179 million Euros, followed by the Premier League (-1.137 million Euros). Ligue 1 (-0.593 million Euros) places the least emphasis on age.

@@

## Universal Importance of Goals and Assists
Goals and assists emerged as the most consistent predictors of market value across leagues. The Premier League shows the strongest relationship between goals and market value, followed by Serie A. This finding reflects the premium placed on offensive contributions, particularly in leagues that emphasize high-scoring games or where attacking players are highly valued in the transfer market. Similarly, assists play a crucial role in Germany’s Bundesliga, likely reflecting the league's focus on team-oriented and high-tempo playing styles. These results confirm prior research, which identifies offensive output as a key driver of player valuations in soccer.

@@

While the findings provide valuable results, the study has several limitations. First, it does not account for external factors such as player injuries, media influence, and agent negotiations, all of which are known to influence market value. Second, the analysis focuses on a single season (2023/24), limiting its ability to capture temporal trends and variations caused by external events such as the COVID-19 pandemic or changes in league regulations. Third, the use of linear regression assumes a straightforward relationship between predictors and market value, which may oversimplify complex interactions or nonlinear effects (e.g., diminishing returns on goals). Finally, the study’s focus on Europe’s top five leagues excludes emerging leagues, such as Major League Soccer (MLS) or the Saudi Pro League, where market dynamics may differ significantly.

@@

The findings of this study have important implications for clubs, agents, and analysts in the soccer industry. Clubs can use these insights to refine their scouting and recruitment strategies by focusing on metrics that are most valued in their specific league, such as goals in the Premier League or assists in the Bundesliga. Agents can leverage the results to highlight the most marketable aspects of their clients during contract negotiations, emphasizing factors like offensive contributions or club prestige. Additionally, analysts and researchers can build on these findings to explore broader trends in player valuations and market dynamics.

Future research could address the limitations of this study by incorporating external variables such as media coverage, player injuries, and sponsorship deals. Expanding the analysis to include data from multiple seasons would provide a more comprehensive understanding of long-term trends and the impact of external shocks, such as economic crises or regulatory changes. Furthermore, the use of advanced machine learning models could capture more complex relationships between predictors and market value, offering a deeper understanding of the factors that drive player valuations. Including emerging leagues and non-European markets would also provide a more global perspective on the determinants of market value, shedding light on how valuations differ between established and developing soccer ecosystems.

@@

## Data Retrieval

This section provides a detailed explanation of how the data for this analysis was collected and processed. The data spans several categories, including market values, performance metrics, national team rankings, and club rankings. All data was meticulously gathered from reliable sources, as described below.

@@

The market value data for players in the five major European leagues was retrieved from [Transfermarkt.com](https://www.transfermarkt.com), a widely recognized platform for football statistics and player valuations. The Python script, located at `scripts/02-scrape_data.py`, was used to scrape data from Transfermarkt. The script leverages the `requests` library for fetching webpage content and `BeautifulSoup` for parsing HTML.

@@

National team rankings were sourced from **[FIFA World Ranking](https://www.fifa.com/fifa-world-ranking/men)**, a globally recognized metric for evaluating the performance of national football teams based on international match results. These rankings provide valuable context for assessing the market values of players, particularly those participating in international competitions or representing highly ranked national teams.

@@

These rankings were manually extracted for each club and merged into the dataset by matching player club affiliations. This method provided an accurate and up-to-date measure of club performance, offering crucial context for analyzing player market values.

>>>>

---
title: "Determinants of Soccer Players' Market Values in Major European League(s)"
subtitle: "Goals, Assists, and Age Dominate, While League-Specific Patterns Add Complexity"
author: 
  - John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Player_Market_Value_Analysis"
date: today
date-format: long
abstract: "The global soccer transfer market is shaped by diverse factors influencing player valuations across leagues. This study analyzes determinants of market value in five major European leagues using data from the 2023/24 season. Goals and assists consistently emerge as the strongest predictors, with the Premier League valuing goals highest and the Bundesliga emphasizing assists, while age negatively affects value across all leagues, particularly in La Liga. These observations highlight universal trends and league-specific valuation dynamics, offering actionable guidance for clubs, agents, and analysts."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The global soccer transfer market is a high-stakes, multi-billion-Euro industry where decisions on player valuations can significantly influence team-building strategies, investment decisions, and negotiations. Player market values are determined by a mix of tangible performance metrics, such as goals scored or assists provided, and intangible factors like potential, reputation, and team context. While certain predictors are universally recognized, such as offensive contributions and age, the relative importance of these factors varies widely across leagues with differing financial capabilities, tactical preferences, and competitive levels. Understanding the dynamics behind these variations is important for clubs, agents, and analysts aiming to optimize their strategies in this complex market.

This study examines the determinants of soccer players’ market values across five major European leagues: Premier League (England), La Liga (Spain), Serie A (Italy), Bundesliga (Germany), and Ligue 1 (France). These leagues not only dominate the global football landscape but also present unique economic and competitive ecosystems, making them ideal for a comparative analysis. Using player data from the 2023/24 season, sourced from platforms such as @citeTransfermarkt, @citeStathead, @citeFIFA, and @citeFootballDatabase, this research addresses a vital gap: how do the drivers of player market value differ across these leagues?

@@

The results show both universal and league-specific patterns in player valuation. Offensive contributions, such as goals and assists, consistently emerged as significant predictors, with goals having the strongest influence in the Premier League and assists being most impactful in the Bundesliga. Age also played an important role, with younger players commanding higher valuations, particularly in La Liga. Contextual variables like club and national team rankings demonstrated notable variation, reflecting the differing emphasis on team prestige across leagues.

This study contributes to the literature on soccer economics by uncovering how market dynamics vary across Europe’s top leagues. For clubs, the findings provide actionable insights for tailoring recruitment and investment strategies. For agents, the results highlight the most marketable player attributes for negotiations. Methodologically, the integration of elevated statistical and programming tools offers a replicable framework for analyzing market dynamics in soccer and other sports.

The paper is structured as follows: @sec-data describes the dataset and variable construction processes, emphasizing the integration of data from multiple sources. @sec-model details the modeling approach and validation techniques. @sec-res presents the empirical results, identifying key differences across leagues. @sec-disc discusses the implications of these findings, highlights the limitations of the study, and outlines future research directions. @sec-appendix provides an in-depth examination of data cleaning methodologies, in-depth documentation of observational data collection processes, datasheets for each dataset, and full model results. By bridging methodological rigor with actionable insights, this study advances our understanding of the determinants of soccer players’ market values.

@@

The dataset was created by merging multiple data sources using player names as unique identifiers. **Transfermarkt** provided player market values in Euros, which serve as the dependent variable in this study. Performance metrics such as goals and assists were sourced from **Stathead**, a platform that aggregates detailed statistics on football players. To ensure consistency and comparability, rankings were applied to national teams and clubs. National team rankings (1 to 210) were extracted from **FIFA**, while club rankings were sourced from **Footballdatabase**, reflecting both domestic and international performances (See @sec-ret for a detailed walkthrough of retrieving data).

After merging these sources, high-level cleaning was conducted to address missing data and standardize variable formats, and the players who are classified as goalkeepers are dropped (See @sec-clean). The final cleaned datasets for each league include seven variables, as outlined below:

@@

The England dataset provides a thorough view of player characteristics in the English Premier League. As illustrated in @fig-emp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower market value brackets (below €50 million). This skewness reflects a superstar effect, where a small subset of players, often international stars or highly sought-after talent, command disproportionately higher valuations.

@fig-efp explores the distributions of key predictor variables, including age, goals, assists, national team ranking, club ranking, and minutes played. These distributions show important patterns in the dataset:

@@

# Reshape data for faceting (excluding market_value)
data_long <- england_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    Age = age,
    Goals = goals,
    Assists = assists,
    `National Team Ranking` = national_team_ranking,
    `Club Ranking` = club_ranking,
    `Minutes Played` = minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "Age",
      "Goals",
      "Assists",
      "National Team Ranking",
      "Club Ranking",
      "Minutes Played"
    ))
  )

@@

# Summarize numeric variables
numeric_summary <- england_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    `Market Value (in Million Euros)` = market_value_million,
    Age = age,
    Goals = goals,
    Assists = assists,
    `Club Ranking` = club_ranking,
    `National Team Ranking` = national_team_ranking,
    Minutes = minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The France dataset offers interpretation into player profiles from Ligue 1, one of Europe’s top football leagues. Similar to other leagues, the distribution of market values in the French dataset, as depicted in @fig-fmp, is heavily right-skewed. Most players have market values below €50 million, with only a few elite players commanding significantly higher valuations. This reflects the trend where top-tier talents, often representing national teams or excelling in international competitions, are valued much higher than their peers.

@@

- **Age**: The age distribution in Ligue 1 peaks around 24 years, with a majority of players between 20 and 30 years old. This highlights the league's balance between developing young talent and retaining experienced professionals. The maximum age in the dataset is 39 years, underscoring the presence of veteran players in the league.
- **Goals and Assists**: The distributions for goals and assists are highly right-skewed, with most players contributing fewer than 5 goals or assists in a season. This indicates that offensive production is concentrated among a small group of players, typically forwards or attacking midfielders.
- **Club Ranking**: Club rankings span a wide range, from 11 to 568, reflecting the diversity in club performance levels within the league.
- **National Team Ranking**: National team rankings also show significant variation, with players representing countries ranked in the top 10 as well as those from lower-ranked national teams. This diversity emphasizes Ligue 1's role as a platform for both elite and emerging talent.
- **Minutes Played**: The distribution of minutes played indicates that a significant proportion of players log substantial game time (over 1,000 minutes), with some exceeding 3,000 minutes, indicating their vital role in team strategies.

@@

@tbl-sf provides a detailed statistical summary of the numeric variables in the dataset. The mean market value stands at €7.98 million, significantly lower than the Premier League, indicating the varying financial dynamics across leagues. The average age is 23.91 years, reinforcing the league's focus on younger talent. Other variables such as goals and assists demonstrate low means relative to their maximum values, reflecting the rarity of exceptional performances.

@@

# Reshape data for faceting and specify the order of variables
data_long <- france_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    Age = age,
    Goals = goals,
    Assists = assists,
    `National Team Ranking` = national_team_ranking,
    `Club Ranking` = club_ranking,
    `Minutes Played` = minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "Age",
      "Goals",
      "Assists",
      "National Team Ranking",
      "Club Ranking",
      "Minutes Played"
    ))
  )

@@

# Summarize numeric variables for France
numeric_summary_france <- france_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    `Market Value (in Million Euros)` = market_value_million,
    Age = age,
    Goals = goals,
    Assists = assists,
    `Club Ranking` = club_ranking,
    `National Team Ranking` = national_team_ranking,
    Minutes = minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

@fig-gfp explores the distributions of predictor variables, highlighting key patterns:

@@

The position distribution, depicted in @fig-gpp, demonstrates that defenders (DF) comprise the largest share of players (31.7%), followed by midfielders (MF) and forward-midfield hybrids (FWMF). Pure forwards (FW) form 9.6% of the players, while hybrid positions like FWDF make smaller contributions. This composition underscores the league's focus on tactical balance, with strong emphasis on midfield and defensive strategies.

@@

# Reshape data for faceting (excluding market_value)
data_long <- germany_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    Age = age,
    Goals = goals,
    Assists = assists,
    `National Team Ranking` = national_team_ranking,
    `Club Ranking` = club_ranking,
    `Minutes Played` = minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "Age",
      "Goals",
      "Assists",
      "National Team Ranking",
      "Club Ranking",
      "Minutes Played"
    ))
  )

@@

# Summarize numeric variables for Germany
numeric_summary_germany <- germany_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    `Market Value (in Million Euros)` = market_value_million,
    Age = age,
    Goals = goals,
    Assists = assists,
    `Club Ranking` = club_ranking,
    `National Team Ranking` = national_team_ranking,
    Minutes = minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The Italy dataset provides an detailed analysis of player characteristics within the Serie A league. As shown in Figure @fig-imp, the market value distribution is right-skewed, with a majority of players concentrated in lower valuation brackets (below €30 million). This distribution is characteristic of the league, where the presence of a few superstar players significantly elevates the upper tail.

@@

# Reshape data for faceting and specify the order of variables
data_long <- italy_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    Age = age,
    Goals = goals,
    Assists = assists,
    `National Team Ranking` = national_team_ranking,
    `Club Ranking` = club_ranking,
    `Minutes Played` = minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "Age",
      "Goals",
      "Assists",
      "National Team Ranking",
      "Club Ranking",
      "Minutes Played"
    ))
  )

@@

# Summarize numeric variables for Italy
numeric_summary_italy <- italy_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    `Market Value (in Million Euros)` = market_value_million,
    Age = age,
    Goals = goals,
    Assists = assists,
    `Club Ranking` = club_ranking,
    `National Team Ranking` = national_team_ranking,
    Minutes = minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The Spain dataset offers a detailed analysis of player characteristics in La Liga. As illustrated in @fig-smp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower valuation brackets (below €30 million). This trend reflects the league's structure, where a few superstar players dominate the upper tail of the distribution, commanding significantly higher valuations.

@fig-sfp shows the distributions of key predictor variables:

@@

@fig-spp illustrates the percentage distribution of player positions. Defenders (DF) comprise the largest group (32.3%), followed by midfielders (MF) at 24.4%. Hybrid roles such as forward-midfield (FWMF) and forward (FW) are well-represented, reflecting La Liga’s tactical diversity and emphasis on attacking football. Smaller proportions of hybrid defensive positions (FWDF, MFDF) further highlight the league's strategic aspect.

@@

# Reshape data for faceting and specify the order of variables
data_long <- spain_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    Age = age,
    Goals = goals,
    Assists = assists,
    `National Team Ranking` = national_team_ranking,
    `Club Ranking` = club_ranking,
    `Minutes Played` = minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "Age",
      "Goals",
      "Assists",
      "National Team Ranking",
      "Club Ranking",
      "Minutes Played"
    ))
  )

@@

# Summarize numeric variables for Spain
numeric_summary_spain <- spain_data %>%
  select(
    market_value_million,
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  rename(
    `Market Value (in Million Euros)` = market_value_million,
    Age = age,
    Goals = goals,
    Assists = assists,
    `Club Ranking` = club_ranking,
    `National Team Ranking` = national_team_ranking,
    Minutes = minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

@@

The construction of variables in this study ensures their relevance and accuracy. Player market values were obtained from Transfermarkt, a widely regarded source for estimating market valuations. These values are based on a combination of player performance, market demand, and subjective expert evaluation, making them a holistic representation of player worth. Performance metrics, such as goals and assists, were sourced from Stathead, a trusted repository of sports data. National team and club rankings were quantified to enable comparison across leagues. National team rankings were sourced from FIFA’s official ranking system, which reflects international match results, strength of opposition, and match importance. Club rankings were sourced from Footballdatabase, which evaluates team performance using a consistent point-based methodology (See @sec-meth for a thorough description of data collection processes)

@@

The selected dataset is uniquely suited for this analysis due to its extensive scope and integration of market, performance, and contextual data. While alternative sources such as Opta or WyScout provide more granular match-level data, their focus on in-game metrics (e.g., pass accuracy, expected goals) lacks the broader valuation context provided by Transfermarkt. Moreover, FIFA rankings and Footballdatabase rankings are well-established benchmarks for quantifying national and club performance, making them preferable over less transparent systems.

@@

In this study, we constructed a linear regression model for each of the five major European football leagues to analyze the determinants of player market value. The dependent variable, market value $(\text{Market Value}_i)$, was scaled to millions of Euros for easier interpretability. The model incorporates a wide-ranging set of predictors, including player performance metrics, contextual factors, and positional variables, to capture the multi-dimensional aspects influencing market valuation.

@@

The models were implemented using the R programming language, specifically utilizing the `lm()` function for linear regression. For each league, data were split into training and testing sets to validate model performance. A stratified train-test split (80% training, 20% testing) was performed using the `caret` package to ensure balanced representation of market value distributions across the sets.

@@

Age negatively affects market value across all leagues, reflecting the premium placed on youth. @fig-ae demonstrates that La Liga (Spain) places the strongest emphasis on younger players, with a coefficient of -1.179 million Euros, followed by the Premier League (-1.137 million Euros). Ligue 1 (-0.593 million Euros) places the least emphasis on age.

@@

## Universal Importance of Goals and Assists
Goals and assists emerged as the most consistent predictors of market value across leagues. The Premier League shows the strongest relationship between goals and market value, followed by Serie A. This finding reflects the premium placed on offensive contributions, particularly in leagues that emphasize high-scoring games or where attacking players are highly valued in the transfer market. Similarly, assists play a significant role in Germany’s Bundesliga, likely reflecting the league's focus on team-oriented and high-tempo playing styles. These results confirm prior research, which identifies offensive output as a key driver of player valuations in soccer.

@@

While the findings provide useful results, the study has several limitations. First, it does not account for external factors such as player injuries, media influence, and agent negotiations, all of which are known to influence market value. Second, the analysis focuses on a single season (2023/24), limiting its ability to capture temporal trends and variations caused by external events such as the COVID-19 pandemic or changes in league regulations. Third, the use of linear regression assumes a straightforward relationship between predictors and market value, which may oversimplify complex interactions or nonlinear effects (e.g., diminishing returns on goals). Finally, the study’s focus on Europe’s top five leagues excludes emerging leagues, such as Major League Soccer (MLS) or the Saudi Pro League, where market dynamics may differ significantly.

@@

The findings of this study have important implications for clubs, agents, and analysts in the soccer industry. Clubs can use these observations to refine their scouting and recruitment strategies by focusing on metrics that are most valued in their specific league, such as goals in the Premier League or assists in the Bundesliga. Agents can utilize the results to highlight the most marketable aspects of their clients during contract negotiations, emphasizing factors like offensive contributions or club prestige. Additionally, analysts and researchers can build on these findings to explore broader trends in player valuations and market dynamics.

Future research could address the limitations of this study by incorporating external variables such as media coverage, player injuries, and sponsorship deals. Expanding the analysis to include data from multiple seasons would provide a more in-depth understanding of long-term trends and the impact of external shocks, such as economic crises or regulatory changes. Furthermore, the use of cutting edge machine learning models could capture more complex relationships between predictors and market value, offering a deeper understanding of the factors that drive player valuations. Including emerging leagues and non-European markets would also provide a more global perspective on the determinants of market value, highlighting how valuations differ between established and developing soccer ecosystems.

@@

## Data Retrieval{#sec-ret}

This section provides a detailed explanation of how the data for this analysis was collected and processed. The data spans several categories, including market values, performance metrics, national team rankings, and club rankings. All data was gathered from reliable sources, as described below.

@@

The market value data for players in the five major European leagues was retrieved from [Transfermarkt.com](https://www.transfermarkt.com), a widely recognized platform for football statistics and player valuations. The Python script, located at `scripts/02-scrape_data.py`, was used to scrape data from Transfermarkt. The script utilizes the `requests` library for fetching webpage content and `BeautifulSoup` for parsing HTML.

@@

National team rankings were sourced from **[FIFA World Ranking](https://www.fifa.com/fifa-world-ranking/men)**, a globally recognized metric for evaluating the performance of national football teams based on international match results. These rankings provide useful context for assessing the market values of players, particularly those participating in international competitions or representing highly ranked national teams.

@@

These rankings were manually extracted for each club and merged into the dataset by matching player club affiliations. This method provided an accurate and up-to-date measure of club performance, offering pivotal context for analyzing player market values.
